Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton

### 2022-06-12 text/Manuscript.Rmd
MS: results section 1 edits

**Nitric oxide synthase is expressed in neurons in the cPRC circuit**

First we performed a nitric oxide synthase (NOS) homology search on Platynereis transcriptome data and identified one NOS gene. We examined the gene expression pattern of NOS using in situ hybridization chain reaction (HCR). In the 2-day-old larvae, the NOS gene was strongly expressed in four cells, one pair on the dorsal side and one pair on the ventral side (Fig. 1A). NOS cells were labelled using 12 kbp upstream of the transcription start point of the NOS gene and NOS cells were identified as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig. 1B). We named these NOS interneurons, INNOS (INNOSdl, dr, vl, vr) (Fig. 1C). The cPRC circuit has been reported as a neural circuit involved in c-opsin1 mediated UV avoidance behaviour (Verasztó et al., 2017). INNOS were presynaptic from the cPRCs and postsynaptic to the RGW interneurons in the cPRC circuit (Fig. 1D). To determine whether NO is involved in UV avoidance behaviour, we generated mutant lines with the CRISPR/Cas-9 system and examined their behavioural responses to UV light.

@@

![**Figure 4. NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (B) The intensity changes in cGMP fluorescence over time for the three conditions are shown. Gray traces show the optical responses to individual intensities (n = 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars). (C) The mean intensity (Fmean/F0) after acting on the solution. Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig4.png){width="100%"}

>>>>

**Nitric oxide synthase is expressed in interneurons of the UV-avoidance circuit**

We identified a single *nitric oxide synthase* (*NOS*) gene in the *Platynereis dumerilii* genome and transcriptome data. Phylogenetic analysis of NOS proteins indicate that the *Platynereis* NOS belongs to an orthology group of metazoan NOS sequences [need to add tree]. To characterise the expression pattern of *NOS* we used in situ hybridization chain reaction (HCR) and transient transgenesis. In two-day-old larvae, we detected *NOS* expression in four cells in the apical organ region (Fig. 1A). The same four cells cells were labelled with a reporter construct driving palmitoylated-tdTomato by a 12 kbp upstream genomic region of the *NOS* gene. This reporter also revealed the axonal projections of the NOS-expressing neurons. The position and axonal projections of the four *NOS*-expressing cells allowed us to identify the same cells as four interneurons in the ciliary-type photoreceptor cells (cPRCs) circuit from Platynereis reconstruction data (Fig. 1B). We named these NOS interneurons, INNOS (INNOSdl, dr, vl, vr) (Fig. 1C). The cPRC circuit has been reported as a neural circuit involved in c-opsin1 mediated UV avoidance behaviour (Verasztó et al., 2017). INNOS were presynaptic from the cPRCs and postsynaptic to the RGW interneurons in the cPRC circuit (Fig. 1D). To determine whether NO is involved in UV avoidance behaviour, we generated mutant lines with the CRISPR/Cas-9 system and examined their behavioural responses to UV light.

@@

![**Figure 4. NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (B) The intensity changes in cGMP fluorescence over time for the three conditions are shown. Gray traces show the optical responses to individual intensities (n = 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars). (C) The mean intensity (Fmean/F0) after acting on the solution. Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Fig6.png){width="100%"}
